---
title: "Analysis Aquatic Lineages"
author: "Nina Pak"
date: "8/30/2021"
output: html_document
---

## R Markdown

```{r}
set.seed(1990)
library(ape)
library(phytools)
library(geiger)
library("RColorBrewer")
library(wesanderson)
#library(phylobase) - maybe for another project
library(corHMM)
#Set Working Directory and Upload file. Pick the correct tree file. 
setwd("/Users/npak/Desktop/E")
tree <- read.nexus("/Users/npak/Desktop/E/D06112019.nex")
#rooting tree
is.rooted(tree)
#no, not yet
outgroup <- c("Merope_tuber", "Panorpa_sp", "Nannochorista_sp", "Ctenocephalides_felis", "Micropteryx_calthella", "Microchorista_philpotti","Deuterophlebia_coloradensis")
rooted_tree <- root(tree, outgroup, resolve.root = TRUE)
is.rooted(rooted_tree)

#tree with modified branches for analysis
dst <- rooted_tree
dst$edge.length[dst$edge.length==0]<-max(nodeHeights(rooted_tree))*1e-6

#load annotated tips
svl<-read.csv("/Users/npak/Desktop/AL folder/test_bloodfeeding.csv", header = TRUE, row.names = 1)
#create a value to set for tip labels
fmode.aquatic<-as.factor(setNames(svl[,1], rownames(svl)))
fmode.blood <-as.factor(setNames(svl[,2], rownames(svl)))
#svl (characterizes as factor like a data frame); 
#Column 1 is coded for Found having an aquatic life history in its larval stage. 
#Column 2 is coded for Found for Blood Feeding as an Adult. Keep track of fmode!
```

## ROOTED VERSION
Using Pagel's method, we can compare independent and dependent models. This shows the rates of change between the states between Aquatic to Hematophagy. The boldness is proportional to the rate (Higher rate is more bolded).
```{r}
library(phytools)
x_dst <- setNames(svl$Overall, rownames(svl))
y_dst <- setNames(svl$Yuval, rownames(svl))

fit_xy_dst <-fitPagel(dst, x_dst, y_dst)
fit_xy_dst
plot(fit_xy_dst, lwd.by.rate=TRUE)
```

```{r}

fit_yx_dst <-fitPagel(dst, y_dst, x_dst)
fit_yx_dst
plot(fit_yx_dst, lwd.by.rate=TRUE)
```

```{r}
fit.x_dst <- fitPagel(dst,x_dst,y_dst, dep.var = "x")
fit.x_dst
plot(fit.x_dst,lwd.by.rate=TRUE)
```

```{r}
fit.y_dst<-fitPagel(dst,x_dst,y_dst,dep.var="y")
fit.y_dst
pdf(file ="dependenty-plot.pdf")
plot(fit.y_dst,lwd.by.rate=TRUE)
```

 We can see this in our weighted AIC and AIC scores. 

```{r}
aic_dst<-setNames(c(fit_xy_dst$independent.AIC,
    fit.x_dst$dependent.AIC,
    fit.y_dst$dependent.AIC,
    fit_xy_dst$dependent.AIC),
    c("independent","dependent x",
    "dependent y","dependent x&y"))
aic_dst
```

Weighted AIC scores. 
```{r}
aic.w(aic_dst)

plot(fit.y_dst,lwd.by.rate=TRUE)
fit.y_dst
fit_xy_dst
```

#Phylo.D

We analyzed D using the ‘phylo.d’ function in the R package caper (Orme et al., 2018), and conducted a significance test by comparing observed D with distributions of simulated values produced under two scenarios: random shuffling of traits on the tree, or simulation of a binary trait under a Brownian motion model. 
```{r}
library(caper)

AquaticFlies <- comparative.data(dst, svl_test, name)
bloodPhyloD <- phylo.d(AquaticFlies, binvar = Yuval)
aquaticPhyloD <- phylo.d(AquaticFlies, binvar = Overall)

print(bloodPhyloD)
plot(bloodPhyloD)

bloodPhyloD$Pval1
bloodPhyloD$Pval0

print(aquaticPhyloD)
plot(aquaticPhyloD)

aquaticPhyloD$Pval1
aquaticPhyloD$Pval0

```

#stochastic character mapping for aquatic life history
```{r}
set.seed(100)
## simulate single stochastic character map using 'equal rates'
mtree_a_er_one <- make.simmap(dst, fmode.aquatic, model = "ER", pi= "estimated", Q = "empirical")

#'all rates different'
mtree_a_ard_one <- make.simmap(dst, fmode.aquatic, model = "ARD", pi= "estimated", Q = "empirical")

#branched tree, now for 100 simulations
mtrees_a_er_100 <-make.simmap(dst, fmode.aquatic, model = "ER", nsim = 100, pi= "estimated", Q = "empirical")
mtrees_a_ard_100 <-make.simmap(dst, fmode.aquatic, model = "ARD", nsim = 100, pi="estimated", Q = "empirical")

#summarize those simulations
pd_er_0_q<-summary(mtrees_a_er_100)
pd_ard_0_q <-summary(mtrees_a_ard_100)

###
#plot this one tree - ER
cols3<-setNames(c("#16A076","#D46127"),levels(fmode.aquatic))
plot(mtree_a_er_one, ftype="off",colors = cols3, fsize=0.4)
nodelabels(pie = pd_er_0_q$ace, piecol =cols3, cex=0.25)
add.simmap.legend(colors=cols3,prompt=FALSE,x=0.5*par()$usr[1],
    y=-25.8*par()$usr[3],fsize=0.8)

cols3<-setNames(c("#16A076","#D46127"),levels(fmode.aquatic))
plot(mtree_a_ard_one,ftype="off",colors = cols3, fsize=0.4)
nodelabels(pie = pd_ard_0_q$ace, piecol =cols3, cex=0.25)
add.simmap.legend(colors=cols3,prompt=FALSE,x=0.5*par()$usr[1],
    y=-25.8*par()$usr[3],fsize=0.8)
```

#stochastic character mapping for blood-feeding 

```{r}
pdf(file="SCM-ER-ARD-FIGURE - h - woQ-1.pdf",width=10,height=10)
fmode.blood <-as.factor(setNames(svl[,2], rownames(svl)))
## simulate single stochastic character map using empirical Bayes method
mtree_h_er_one <- make.simmap(dst, fmode.blood, model = "ER", pi= "estimated", Q = "empirical")
## simulate single stochastic character map using empirical Bayes method
mtree_h_ard_one <- make.simmap(dst, fmode.blood, model = "ARD", pi= "estimated", Q = "empirical")
#branched tree, now for 100 simulations
mtrees_h_er_100 <-make.simmap(dst, fmode.blood, model = "ER", nsim = 100, pi= "estimated", Q = "empirical")
mtrees_h_ard_100 <-make.simmap(dst, fmode.blood, model = "ARD", nsim = 100, pi="estimated", Q = "empirical")

#summarize those simulations
pd_er_0_h<-summary(mtrees_h_er_100)
#pd_er_0_h
pd_ard_0_h <-summary(mtrees_h_ard_100)
#pd_ard_0_h

#plot results
cols2<-setNames(c("black","red"),c("Other", "Aquatic"))
plot(mtree_h_er_one,ftype="off")
cols<-setNames(c("black","red"),c("Other", "Aquatic"))
nodelabels(pie = pd_er_0_h$ace, piecol =cols2, cex=0.25)
add.simmap.legend(colors=cols,prompt=FALSE,x=0.5*par()$usr[1],
    y=-25.8*par()$usr[3],fsize=0.8)

plot(mtree_h_ard_one,ftype="off")
cols<-setNames(c("black","red"),c( "Other","Aquatic"))
nodelabels(pie = pd_ard_0_h$ace, piecol =cols2, cex=0.25)
add.simmap.legend(colors=cols,prompt=FALSE,x=0.5*par()$usr[1],
    y=-25.8*par()$usr[3],fsize=0.8)

```

